---
title: "ExploratoryDataAnalysis"
author: "Rajesh Jagannath"
date: "August 3, 2016"
output: github_document
---

### 1. Introduction
I am using Datasets provided by Cincinnati Area Geographic Information System (CAGIS).  A Street car is being introduced in the City of Cincinnati. Its economic benefit is being analyzed and forecast in this Project. A Buffer zone has been established around the street-car route. It is further divided into CORE, CENTER and EDGEA subset of features  have been seleced from the original .csv file. In the following section we explore the data in 3 csv files by plotting the market Land Value  and a count of onbservations catergorized by Existing Land Use Code.

### 2. Pre-Procesing 
This program  reads in the CENTER, EDGE and CORE csv files
```{r, warning=FALSE, message= FALSE}
library(dplyr)
library(RSQLite)
library(tidyr)
library(ggplot2)
library(readr)
library(stringr)
library(scales)
library(mixtools)
library(readxl)
library(ggmap)
```

### 3. Data Loading
```{r, echo = FALSE}
setwd("/Users/rajesh/Desktop/Coursera/SpringBoardGithub/StreetCar0719")
```
```{r, message=FALSE, warning=FALSE, results='hide'}
# read in the csv 
stcar_center <- read_csv("./streetcarbuffer_parcels/parcel_csv_050616/StreetCarParcels_CENTER.csv")
stcar_core   <- read_csv("./streetcarbuffer_parcels/parcel_csv_050616/StreetCarParcels_CORE.csv")
stcar_edge   <- read_csv("./streetcarbuffer_parcels/parcel_csv_050616/StreetCarParcels_EDGE.csv")
```
### 4. Prepare data 
Subset some interesting features - PARCELID, Existing Land Use Code, Mkt value of the land/ improvements, Sale Amt., Area  and Acres
```{r }
selected_var_v <- c("PARCELID", "EXLUCODE", "MKTLND", "MKTIMP", "MKT_TOTAL_","ADDRNO","ADDRST", "ADDRSF")


parcelid_center<- stcar_center[selected_var_v]
parcelid_edge  <- stcar_core[selected_var_v]
parcelid_core  <- stcar_edge[selected_var_v]

# Existing land use code  description
exlu_desc <- c("?- Unknown", "C - Commercial", "ED - Educational ", "HI - heavy Industrial", "IN - Instituitional ", "LI - Light Industrial" ,"MF - Multi-Family", "MU - Mixed use", "O - Office", "PR - Parks and Recreation ", "PS - Public / Semi-Public", "PU - Public utilities", "SF - Single Family", "TF - Two Family Dwelling ", "VA - Vacant")
```
###5a. Exploratory Analysis on Dataset - CORE
```{r}
str(parcelid_core)
```
Total number of parcels and the mean Market value of the Land  in CORE
```{r, echo = FALSE}
print(paste("Total number of parcels in CORE is         ", nrow(parcelid_core)))
print(paste("The mean Market value of the land in CORE  ", mean(parcelid_core$MKTLND)))
print(paste("The count of parcels with 0 MKT value is    ", length(which(parcelid_core$MKTLND == 0))))
```



#### Plot- CORE 
Most High Value properties are  in the Core buffer area are C- Commercial, VA - , PS - and O- Office
We need to remove or impute the lone data with LandUse classified as "?"
One office building has Mkt Land value of 25Million


```{r coreplots, echo=FALSE}
 
# Data Layer

  n <-(ggplot)
# 
# # Data + Aesthetics Mapping 

# n <-ggplot(data,aes(x= data$Department.Title, y = data$Total.Payments, color=data$Job.Class.Title, alpha=0.4, size=data$Overtime.Pay))

n <- ggplot(parcelid_core,aes(x = factor(parcelid_core$EXLUCODE), y=(parcelid_core$MKTLND/1000), color = factor(parcelid_core$EXLUCODE)))
  
# 
# # Data + aes + Geometries 
# n <- n + geom_jitter() 
n <- n + geom_jitter()
#
# # Data + Aesthetic Mapping +  Geom + Facets
# #n <- n + facet_grid(. ~ Department.Title ) 
# 
# #Data + Aesthetics + Geoms + Facets + Statistics
# # ... add a linear regression model here
# 
# #Data + Aesthetics + Geoms + Facets + Statistics + Co-ordinates
# #n <- n + scale_y_discrete(limits=c("00000","120000"), breaks=seq(00000,120000,10000))
 n <- n + xlab("Existing Land Use Code CORE BUFFER ZONE")
 n <- n + ylab("Land Market value in 000s")
 n <- n + scale_color_discrete(labels=paste(exlu_desc))

# #Data + Aesthetics + Geoms + Facets + Statistics + Co-ordinates + Theme 
# n <- n + theme(axis.text.x = element_text(angle=305))

 print(n)
 
 m <-(ggplot)
 # Data + aes 
 m <- ggplot(parcelid_core, aes(x= factor(parcelid_core$EXLUCODE),fill=factor(parcelid_core$EXLUCODE)))
 # Data + aes + Geom 
 m <- m + geom_bar()
 
 # Data + aes + geom + facets + stats + Co-ords
 
 m <- m + xlab("Existing Land Use Code ")
 m <- m + scale_fill_discrete(labels=exlu_desc)
 m <- m + guides(fill=guide_legend(reverse=TRUE))

# 
# Theme
m <- m + coord_flip()
 
m <- m + geom_text(stat="count", color="white", hjust=1.0, size=3,
            aes(y=..count.., label=..count..))
 print(m)
 
#####
 
```

##5b.Exploratory Analysis on Dataset - CENTER
Total number of parcels and the mean Market value of the Land  in CENTER
```{r, echo = FALSE}
str(parcelid_center)
print(paste("Total number of parcels in CENTER is                 ", nrow(parcelid_center)))
print(paste("The mean Market value of the land in ENTER           ", mean(parcelid_center$MKTLND)))
print(paste("The count of parcels in CENTER  with 0 MKT value is  ", length(which(parcelid_center$MKTLND == 0))))

```
  



#### Plot - CENTER

```{r, echo = FALSE}

exlu_desc <- c("C - Commercial", "ED - Educational ", "HI - heavy Industrial", "IN - Instituitional ", "LI - Light Industrial" ,"MF - Multi-Family", "MU - Mixed use", "O - Office", "PR - Parks and Recreation ", "PS - Public / Semi-Public", "PU - Public utilities", "SF - Single Family", "TF - Two Family Dwelling ", "VA - Vacant")


## 6b Plots for CENTER
#  Data Layer

  n <-(ggplot)
# 
# # Data + Aesthetics Mapping 

# n <-ggplot(data,aes(x= data$Department.Title, y = data$Total.Payments, color=data$Job.Class.Title, alpha=0.4, size=data$Overtime.Pay))

n <- ggplot(parcelid_center,aes(x = factor(parcelid_center$EXLUCODE), y=(parcelid_center$MKTLND/1000), color = factor(parcelid_center$EXLUCODE)))
  
# 
# # Data + aes + Geometries 
# n <- n + geom_jitter() 
n <- n + geom_jitter()
#
# # Data + Aesthetic Mapping +  Geom + Facets
# #n <- n + facet_grid(. ~ Department.Title ) 
# 
# #Data + Aesthetics + Geoms + Facets + Statistics
# # ... add a linear regression model here
# 
# #Data + Aesthetics + Geoms + Facets + Statistics + Co-ordinates
# #n <- n + scale_y_discrete(limits=c("00000","120000"), breaks=seq(00000,120000,10000))
 n <- n + xlab("Existing Land Use Code in CENTER BUFFER ZONE")
 n <- n + ylab("Land Market value in 000s")
 n <- n + scale_color_discrete(labels=paste(exlu_desc))

 n <- n + scale_fill_discrete(labels=exlu_desc[-1])
 n <- n + guides(fill=guide_legend(reverse=TRUE))
 
# 
# #Data + Aesthetics + Geoms + Facets + Statistics + Co-ordinates + Theme 
# n <- n + theme(axis.text.x = element_text(angle=305))
# 
 print(n)
 # Explore - How many NA's are there 
 
 m <-(ggplot)
 # Data + aes 
 m <- ggplot(parcelid_center, aes(x=factor(parcelid_center$EXLUCODE),fill=factor(parcelid_center$EXLUCODE)))
 # Data + aes + Geom 
 m <- m + geom_bar()
 # Data + aes + geom + facets + stats + Co-ords
 m <- m + xlab("Existing Land Use Code ")
 m <- m + scale_fill_discrete(labels=exlu_desc[-1])
 m <- m + guides(fill=guide_legend(reverse=TRUE))
# 
# Theme
m <- m + coord_flip()

# Label the bars
m <- m + geom_text(stat="count", color="white", hjust=1.0, size=3,
            aes(y=..count.., label=..count..))
 
 print(m)
 
 
```

##5c. Exploratory Analysis on Dataset - EDGE
```{r, echo = FALSE}
str(parcelid_edge)
```
#### Total records in EDGE
```{r, echo = FALSE}
print(paste("Total number of parcels in EDGE is           ", nrow(parcelid_edge)))
print(paste("The mean  Market value of the land in EDGE is", mean(parcelid_edge$MKTLND)))
print(paste("The count of parcels with 0 MKTLND value is  ", length(which(parcelid_edge$MKTLND == 0))))

```

#### Plot - EDGE

```{r, echo = FALSE}

 exlu_desc <- c("C - Commercial", "ED - Educational ", "HI - Heavy Industrial", "IN - Instituitional ", "LI - Light Industrial" ,"MF - Multi-Family", "MU - Mixed use", "O - Office", "PR - Parks and Recreation ", "PS - Public / Semi-Public", "PU - Public utilities", "SF - Single Family", "TF - Two Family Dwelling ", "VA - Vacant")



### 6c . Including Plots from EDGE
# Data Layer

  n <-(ggplot)
# 
# # Data + Aesthetics Mapping 

# n <-ggplot(data,aes(x= data$Department.Title, y = data$Total.Payments, color=data$Job.Class.Title, alpha=0.4, size=data$Overtime.Pay))

n <- ggplot(parcelid_edge,aes(x = factor(parcelid_edge$EXLUCODE), y=(parcelid_edge$MKTLND/1000), color = factor(parcelid_edge$EXLUCODE)))
  
# 
# # Data + aes + Geometries 
n <- n + geom_jitter() 
#
# # Data + Aesthetic Mapping +  Geom + Facets
 #n <- n + facet_grid(. ~ EXLUCODE ) 
# 
# #Data + Aesthetics + Geoms + Facets + Statistics
# # ... add a linear regression model here
# 
# #Data + Aesthetics + Geoms + Facets + Statistics + Co-ordinates
# #n <- n + scale_y_discrete(limits=c("00000","120000"), breaks=seq(00000,120000,10000))
 n <- n + xlab("Existing Land Use Code in EDGE BUFFER ZONE")
 n <- n + ylab("Land Market value in 000s")
 n <- n + scale_color_discrete(labels=paste(exlu_desc))

 
# 
# #Data + Aesthetics + Geoms + Facets + Statistics + Co-ordinates + Theme 
# n <- n + theme(axis.text.x = element_text(angle=305))
# 
 print(n)

 m <-(ggplot)
 # Data + aes 
 m <- ggplot(parcelid_edge, aes(x= factor(parcelid_edge$EXLUCODE),fill=factor(parcelid_edge$EXLUCODE)))
 # Data + aes + Geom 
 m <- m + geom_bar()
 # Data + aes + geom + facets + stats + Co-ords
 m <- m + xlab("Existing Land Use Code - EDGE Buffer Zone")
 m <- m + scale_fill_discrete(labels=exlu_desc)
 m <- m + guides(fill=guide_legend(reverse=TRUE))
 m <- m + coord_flip()
 
# 
  # Theme
 
 
 m <- m + geom_text(stat="count", color="white", hjust=1.0, size=3,
            aes(y=..count.., label=..count..))
 
 print(m)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# 6. Conclusion
* Exploratory analysis of the three buffer zones on the basis of Existing Land Use classification was done
     + Distribution of Parcels  vs. Existing Land Use
     + Market Value of the parcels on the basis of Existing Land Use
 
* The distribution across the Existing Land Use classes is not uniform


* Further analysis( Exploratory / Forecast ) indicates parcels with High Mrket Land Value are classified as
     + C  - Commercial
     + IN - Instituitional 
     + MU - Mixed Use
     + MF - Multi- Family
     + O  - Office
     + PS - Public / Semi Public
     + SF - Single Family
     + VA - Vacant 
* Parcels with high Market Land values also imply more Property Tax revenue coming into the City
   
   
