---
title: "README.md"
author: "Rajesh Jagannath"
date: "August 2, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Streetcar0719
Steetcar0719 is a directory placing the taxinfo for the capstone project

### OBJECTIVE: 
Determine the economic impact of the introduction of a Street Car route in the City of Cincinnati. The data is from CAGIS and OpenData Cincinnati.

### DATA-SETS :

* CagisOpenDataQuaterly has  shapefiles (.shx) files. The Address.shx would be the most useful one for rendering on  a map.
* Data2 is the primary data-set. These contain 10 years worth of Property Taxes from the assessors office. These are CSV files that have been Extracted from 'fixed-width-format' and coverted to .csv file format. Each directory has the file format described in a pdf file.

* treetcarbuffer_parcels has two subdirs
  + gis_shape_files/gis_shapeFiles_studyarea_Parcels : Used to mark the streetcars parcels
  + old_csv files                   - they are old in the sense that these where first used for exploratory analysis and many records with no addresses were found 
* TrialCode :                      Has first R files with which we explored the data 
  + Read_Position_delimited_File.R  converts fwf to .csv file
  + ReadShapeFile.R                 Read in streetcarbuffer_parcels/gis_shape_files/gis_shapeFiles_studyarea_Parcels/*_PARCELS.shx shape files and rasterizes them into a plot
  + Map_HeatMap_PropVal.R           Exploratory Data Analysis to create a density plot based on csv files( old_csv_files were used
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
